---
title: "Clean Airbnb dataset
output: html_document
---

Some of the code here comes from [this file](https://github.com/alexsingleton/urban_analytics/blob/0bc0f23/04_Descriptive_statistics/P4_Descriptive_Statistics.Rmd),
[licensed with CC-BY](https://github.com/alexsingleton/urban_analytics/blob/0bc0f23/README.md).

We import some [Airbnb](http://insideairbnb.com/get-the-data.html) data for
Amsterdam that was collated in February 2021.

```{r}
# Read in CSV from URL
amsterdam_full <- read.csv("original/listings.csv")
```

This contains a wide array of variables:

```{r}
colnames(amsterdam_full)
```

However, for the example dataset we will subset these data to a limited set of
variables:

```{r}
# Take a subset of the columns.
amsterdam <- subset(amsterdam_full,
                    select = c("id","neighbourhood_cleansed","latitude","longitude","property_type","room_type","bedrooms","price","number_of_reviews"))

# Clean the price data to remove $ and , then convert to numeric (If you are
# doing this outside of the US / UK and have a non English locale set, try
# running Sys.setlocale("LC_ALL","English") before using gsub)
amsterdam$price <- gsub("$","",amsterdam$price,fixed = TRUE)
amsterdam$price <- gsub(",","",amsterdam$price,fixed = TRUE)
amsterdam$price <- as.numeric(as.character(amsterdam$price))

# Remove any records that are not complete
amsterdam <- amsterdam[complete.cases(amsterdam),]

# Show the top six rows
head(amsterdam)
```

In fact, we will start by restricting ourselves to the rooms, houses and
apartments in this dataset.

```{r}
# Give the column data a new name, as a shortcut.
p_type <- amsterdam$property_type
houses_apartments <- subset(amsterdam,
                            p_type == "Private room in house" |
                            p_type == "Private room in apartment" |
                            p_type == "Entire house" |
                            p_type == "Entire apartment")
head(houses_apartments)
```

Save this subset for further analysis.

```{r}
write.csv(houses_apartments, file='processed/amsterdam_airbnb_2_2021.csv')
```
